---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Samuel Hutt (12379926)"
date: "`r Sys.Date()`"
output: distill::distill_article
editor_options: 
  markdown: 
    wrap: 72
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers
    (Martrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your
    results.
-   Use a level 2 heading for each new exercise and answer each subtask
    next to it's bullet point or use a new level 3 heading if you want.
-   Always render the R code for your solutions and make sure to include
    the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless
        specifically instructed to).
-   Always submit both the rendered document(s) as well as your source
    Rmarkdown document. Submit the files separately on moodle, **not**
    as a zip archive.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `2024-exeRcise-sheet-3`. If you forgot how to do this, you can
    follow this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).

    Starting a new R project with git support, called
    "2024-exeRcise-sheet-3".

b)  Commit the files generated by Rstudio.

    Commiting the files ".gitignore" and "2024-exeRcise-sheet-3.Rproj"
    generated by Rstudio.

c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

```         
Committing my changes after finishing each subtask.
```

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)

    **Two strengths of git:**

    -   The history lets you review changes and revert them if issues
        arise.
    -   It's very difficult to lose files with Git.

    **Two weaknesses of git:**

    -   History takes up some file space (but only a small amount).
    -   Has difficulties with binary (i.e., non-text) files, such as PDF or Word files.

e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves.
    
    Knitting this exercise sheet and ignoring the automatically generated HTML page.

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
    
    Creating a new repository on GitHub named "exeRcise-sheet-3" and adding the link to the repository below:
    
    Link to the repository: https://github.com/s-hutt/exeRcise-sheet-3
    
b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
    
    Pushing the code to the new repository by copying and executing the snippet on github.
    
c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.
    
    Pushing regularly the latest changes to GitHub, especially when this sheet is finished.

## Exercise 3: Baby-Names in Munich (3.5 Points)

Download the latest open datasets on given names ("Vornamen") from the
open data repository of the city of Munich for the years `2023` and
`2022`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

a)  Download the data for both years and track it in git. For small
    datasets like these adding them to git is not a problem.
    
    Downloading the datasets and tracking them into git.

b)  Load the data for both years into R. Check the type of the count
    variable ("Anzahl") and look into the data to determine why it is
    not numeric? Fix the problem in an appropriate manner, it is OK if
    some of the counts are inaccurate because of this. Explain your
    solution and the repercussions.
    
```{r}
# Loading the tidyverse:
library(tidyverse)
# Loading the data for the year 2022 into R:
baby_names_munich_22 <- read_csv("data/open_data_portal_2022.csv")
# Loading the data for the year 2023 into R:
baby_names_munich_23 <- read_csv("data/vornamen-muenchen-2023.csv")
# Checking the type of the count variable for 2022:
typeof(baby_names_munich_22$Anzahl)
# The type of the variable "Anzahl" for 2022 is "character".
# Checking the type of the count variable for 2023:
typeof(baby_names_munich_23$Anzahl)
# The type of the variable "Anzahl" for 2023 is "character".
# Looking into the data with the View()-function:
# View(baby_names_munich_22)    - Commented out due to rendering
# View(baby_names_munich_23)    - Commented out due to rendering
```

Answer: The variable "Anzahl" is not numeric because the datasets use the value "4 oder weniger" ("4 or fewer") when the count of the name is not higher than 4.

```{r}
# Fixing the problem:
# Creating a function that cleans an converts the variable "Anzahl" to a numeric variable:
clean_count <- function(df) {
  df %>% 
    mutate(Anzahl = if_else(Anzahl == "4 oder weniger", "4",Anzahl)) %>% 
    mutate(Anzahl = as.numeric(Anzahl))
}
# Applying the new function to both datasets:
baby_names_munich_22 <- clean_count(baby_names_munich_22)
baby_names_munich_23 <- clean_count(baby_names_munich_23)
# Checking the type of the count variable for 2022:
typeof(baby_names_munich_22$Anzahl)
# The type of the variable "Anzahl" for 2022 is "double".
# Checking the type of the count variable for 2023:
typeof(baby_names_munich_23$Anzahl)
# The type of the variable "Anzahl" for 2023 is "double".

# Printing the first 10 rows of both new datasets:
head(baby_names_munich_22)
head(baby_names_munich_23)
```

Explanation of the solution: We addressed the issue by replacing instances of "4 oder weniger" with the numeric value 4 and then converting the "Anzahl" variable to a numeric type.

Explanation of the repercussions: The variable "Anzahl" now consists of numerical values. However, due to the approximation of "4 oder weniger" to 4, it occurs that names appearing less than 4 times are now indicated with the value 4.

c)  Calculate the total number of babies born in Munich in 2022
    and 2023. Which year had the bigger baby-boom?
    
```{r}
# Calculating the total number of babies born in 2022:
total_babies_2022 <- sum(baby_names_munich_22$Anzahl)
# Printing the result:
print(total_babies_2022)
# There is a total of 25199 babies born in 2022.
# Calculating the total number of babies born in 2023:
total_babies_2023 <- sum(baby_names_munich_23$Anzahl)
# Printing the result:
print(total_babies_2023)
# There is a total of 24229 babies born in 2023.
# Comparing the total number of babies born in each year:
if (total_babies_2022 > total_babies_2023) {
  cat("In 2022, there were more babies born in Munich.")
} else if (total_babies_2022 < total_babies_2023) {
  cat("In 2023, there were more babies born in Munich.")
} else {
  cat("The number of babies born in Munich was the same in both 2022 and 2023.")
}
# Calculating the difference:
total_babies_difference <- total_babies_2022 - total_babies_2023
# Printing the solution: 
print(paste("The year 2022 had the bigger baby boom, with a total of", total_babies_difference, "more babies born."))
```
    

d)  Add a new column `year` to both datasets which holds the correct
    year for each.

e)  Combine both datasets into one using `bind_rows()`.

f)  Combine the counts for same names to determine the most popular
    names across both years. Print out the top 10 names in a nicely
    formatted table for both years. Include a table caption.

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from
[Our World in Data](https://ourworldindata.org/) and analyze it, while
determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets>
    and choose a dataset that interests you. You can have a look at
    <https://ourworldindata.org/> to gather some inspiration.
b)  Download the dataset and track it in git.
c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: ...
-   Link: <https://github.com/owid/owid-datasets/>...

d)  Come up with a (research) question you want to answer with the data
    and briefly explain why you believe this is an interesting question
    within one sentence. It should be a question that can be answered
    with the dataset and using R.
e)  Use R to answer your chosen question.
f)  Create a meaningful plot / figure with the dataset. Make sure to
    provide a figure caption (via the chunk options / Rmarkdown) and
    correctly label the figure.

## Final Note

Make sure to push all your commits and changes to GitHub before
submittining the exercise sheet.
